---
title: "Lab 8:"
author: 'Group Member Names - here'
output: html_document
---
Turn in one copy for each group. If group members are not present in class they will be required to complete their own lab to receive credit. Please turn in **both a DOC or PDF file and your R Markdown script**. 

## Lab Overview
For this question, a subset of the tables contained in the History of Baseball database are available. Additional details are available here: [https://www.kaggle.com/seanlahman/the-history-of-baseball](https://www.kaggle.com/seanlahman/the-history-of-baseball). The following tables will be used for these questions:

- player
- all_star
- salary

```{r, message = F}
library(readr)
library(dplyr)
library(knitr)
library(tidyr)
player <- read_csv("http://math.montana.edu/ahoegh/teaching/stat408/datasets/player.csv")
all_star <- read_csv("http://math.montana.edu/ahoegh/teaching/stat408/datasets/all_star.csv")
salary <- read_csv("http://math.montana.edu/ahoegh/teaching/stat408/datasets/salary.csv")

```

### 1. (2 points)
How many players were born in Montana.

```{r}
num_MT <- player %>% filter(birth_state == 'MT') %>% tally() %>% pull()

```

There are `r num_MT` players in the database that were born in Montana.

### 2. (4 points)
Print a table that contains each player born in Montana. The table should contain the player_id as well as given name and their total salary across all years. If salary is not available (pre-1985), include the player but have an NA for salary.

```{r}

total_salary <- salary %>% group_by(player_id) %>% summarize( total_salary = sum(salary))

MT_player <- player %>% filter(birth_state == 'MT') %>% left_join(total_salary, by = "player_id") %>% select(player_id, name_first, name_last, total_salary)

kable(MT_player)
```


### 3. (2 points)

Create a thin dataset for that contains the yearly salaries of David Ortiz, Derek Jeter, and Troy Tulowitzki.

```{r}
thin_salaries <- player %>% filter(name_last %in% c('Tulowitzki', 'Jeter', 'Ortiz')) %>% filter(name_first %in% c('Troy','Derek','David')) %>% left_join(salary, by = 'player_id') %>% select(name_first, name_last, year, salary ) %>% arrange(year)

kable(thin_salaries)
```


### 4. (4 points)

Create a wide dataset for that contains the salaries of David Ortiz, Derek Jeter, and Troy Tulowitzki.

```{r}
wide_salaries <- thin_salaries %>% mutate(year = paste('year',year, sep = '')) %>%
  pivot_wider(names_from = c(year), values_from = salary) 
wide_salaries
```


### 5. (4 points)
Which player(s) made the most appearances as an all star representing the National League (NL)?

```{r}
all_star %>% filter(league_id == 'NL') %>% group_by(player_id) %>% tally() %>% arrange(desc(n)) %>% slice(1:3) %>% left_join(player, by = "player_id") %>% select(name_first, name_last, num_allstars = n) %>% kable()
```

